A component-based DSL for multi-stencil programs

### 2015-07-21 WOLFHPC15/article/concept.tex
wolfhpc15 21_07

%----------------------------------------
\subsection{Multi-stencil programs}
\label{sect:multistencil}
%----------------------------------------
\begin{itemize}
\item textual introduction on the problem, what is a stencil + example. What is a real simulation (more than one stencil and additionnal computations). 
\item Existing solutions with prgramming freedom (instead of PETSc and Trilinos) = stencil compilers which give optimized code for one stencil code. Thus, for distributed memory and HPC, still have to use MPI between optimized codes to get an overall parallelization of the simulation.
\item Short definition of a multi-stencil program and its computations (cite research report for more details)
\end{itemize}

%----------------------------------------
\subsection{Parallelization techniques}
\label{sect:parallel}
%----------------------------------------
\begin{itemize}
\item data parallelism
\item task parallelism
\item hybrid parallelism
\end{itemize}

>>>>

%----------------------------------------
\subsection{Multi-stencil programs}
\label{sect:multistencil}
%----------------------------------------
To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization). In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain (the mesh). Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.

If $x$ represents an element of the mesh, if $N(x)$ is a neighborhood shape around the element $x$, and if $u_t(x)$ is the data (or quantity) to write. Moreover if $V_t(x)$ is the set of data to simulate such that $V_t(x) \cap u_t(x) = \emptyset$, a stencil computation is of the following form

\begin{equation} 
\{u_{t-1}(x), u_{t-1}(N(x)), V_t(x), V_t(N(x)\} \longmapsto u_{t}(x).
\label{space}
\end{equation}

In other words, in a stencil computation the quantity written can be computed using the same quantity at a previous time iteration at $x$ and at $N(x)$. And, any other quantity can also be used in the computation at $x$ and $N(x)$.

A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data. Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations. In this paper we target this kind of complex simulations with more than one stencil computation and also additionnal auxiliary computations, that we call \emph{multi-stencil programs}. The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code~\cite{citeulike12258902, spaaTangCKLL11, Giles2011}. As it will be detailed in the related work, this work is complementary to most stencil compilers and implicit parallelism languages.

\textbf{Do I need the following definitions?}
\begin{mydef}
A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$,
\end{mydef}
where $R$ is the set of data read by the computation, $w$ is the data written by the computation, \textit{exp} is the numerical expression to compute $w$, and $\mathcal{N}$ is a neighborhood shape of the stencil computation.

\begin{mydef}
A local computation is a triplet $l(R,w,\text{exp})$, where $exp$ does not involve a neighborhood function $\mathcal{N}$.
\end{mydef}

%----------------------------------------
\subsection{Parallelization techniques}
\label{sect:parallel}
%----------------------------------------
Three parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.

\paragraph{data parallelism} explain data decomposition and distribution, with a single program to apply on each processor

\paragraph{task parallelism} explain task dag

\paragraph{hybrid parallelism} explain the combination

### 2015-07-21 WOLFHPC15/article/ms_msc.tex
wolfhpc15 21_07

\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} \textbf{(PB: to explain !!!)} is performed on the simulation. A synchronization is needed each time a data read by a stencil computation has been written by a previous computation different from a boundary computation. Actually, as a stencil computation needs neighborhood values, in a domain decomposition values computed by different processors or cores are needed.

>>>>

\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} is performed on the simulation. A synchronization is needed each time a data read by a stencil computation has been written by a previous computation different from a boundary computation. Actually, as a stencil computation needs neighborhood values, in a domain decomposition values computed by different processors or cores are needed.
